---
output:
  html_document: default
  pdf_document: default
---

### Loading packages
```{r, warning=FALSE, message=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(randomForest)
library(e1071)
```

### Downloading and importing data
```{r}
trainurl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testurl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainfn <- "./data/pml-training.csv"
testfn  <- "./data/pml-testing.csv"
download.file(trainurl, destfile=trainfn, method="curl")
download.file(testurl, destfile=testfn, method="curl")
trainDF <- read.csv(trainfn, na.strings=c("NA","#DIV/0!", ""))
testDF <- read.csv(testfn, na.strings=c("NA","#DIV/0!", "")) #validation
```

### Cleaning data
There is a large quantity of empty elements and irrelevant columns.  They are discarded with this process.  
```{r}
trainDF1 <-trainDF[, colSums(is.na(trainDF)) == 0]
testDF1 <-testDF[, colSums(is.na(testDF)) == 0]

removeIdx <- grep("timestamp|^X|window", names(trainDF1))
trainDF2 <- trainDF1[,-c(1, removeIdx )]
testDF2 <- testDF1[,-c(1, removeIdx)]
```

### Partitioning data for cross-validation
Partitioning the data with 70% in the training set for cross-validation
```{r}
classe <- trainDF2$classe
inTrain <- createDataPartition(trainDF2$classe, p=0.7, list=FALSE)
train <- trainDF2[inTrain, ]
test <- trainDF2[-inTrain, ]
```
### Exploring data
The final number of rows and columns for the training and testing sets.  The dataset is highly dimensional with 54 fields.
```{r}
dim(train)
dim(test)
```
Looking at correlations:
```{r}
library(corrplot)
nn <- c(1,ncol(trainDF2)) # excluding name and classe
corrMat <- cor(trainDF2[, -nn])
corr <- findCorrelation(corrMat)
corrplot(corrMat, 
         method="color",
         type="lower",
         order="hclust",
         tl.srt = 45,
         diag = FALSE,
         tl.col = 'black')
```
### Modeling

Starting with a Decision Tree model to assist with model interpretability
```{r}
set.seed(123)
modDT <- rpart(classe ~ ., data=train, method="class")
prp(modDT)

predictDT <- predict(modDT,test,type='class')
cmtree<- confusionMatrix(predictDT, as.factor(test$classe))
cmtree
```

Modeling with Random Forest for improved accuracy performance.
```{r}
controlRF <- trainControl(method="cv", 5)
modelRF <- train(classe ~ ., data=train, method="rf", trControl=controlRF, ntree=100)
modelRF

predictRF <- predict(modelRF, test[, -length(names(test))])
```


